### 2014-06-22 dicook/possessions.Rmd
reflect number of males/females measured

```{r gendermath, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
student2012.sub <- student2012[, c(1, 12, 501)]
colnames(student2012.sub)[1] <- "name"
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub.summary <- summarise(group_by(student2012.sub, name), male=mean(PV1MATH[ST04Q01=="Male"], na.rm=T), 
          female=mean(PV1MATH[ST04Q01=="Female"], na.rm=T))
student2012.sub.summary.m <- melt(student2012.sub.summary)
colnames(student2012.sub.summary.m)[2] <- "gender"
colnames(student2012.sub.summary.m)[3] <- "math"
qplot(gender, math, data=student2012.sub.summary.m, geom="boxplot") + theme(aspect.ratio=1)
# Overall there is a gender gap in math, perhaps 10 points, and the high scores are almost exclusively males
student2012.sub.summary.m$name <- factor(student2012.sub.summary.m$name, 
      levels=student2012.sub.summary$name[order(student2012.sub.summary$male)])
student2012.sub.summary.m$gender <- factor(student2012.sub.summary.m$gender, levels=c("female", "male"))
qplot(name, math, data=student2012.sub.summary.m, colour=gender) + coord_flip()
# Some countries have no gender gap, a few surprises females better than males
student2012.sub.summary.gap <- summarise(group_by(student2012.sub.summary, name), gap=male-female)
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$gap)])
qplot(name, gap, data=student2012.sub.summary.gap, ylim=c(-30, 30)) + geom_hline(yintercept=0, colour="grey80") + coord_flip()
# Gap is interesting
```

>>>>

```{r gendermath, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
student2012.sub <- student2012[, c(1, 12, 501)]
colnames(student2012.sub)[1] <- "name"
student2012.sub$PV1MATH <- as.numeric(student2012.sub$PV1MATH)
student2012.sub.summary1 <- summarise(group_by(student2012.sub, name), male=mean(PV1MATH[ST04Q01=="Male"], na.rm=T), 
          female=mean(PV1MATH[ST04Q01=="Female"], na.rm=T))
student2012.sub.summary.m1 <- melt(student2012.sub.summary1)
colnames(student2012.sub.summary.m1)[2] <- "gender"
colnames(student2012.sub.summary.m1)[3] <- "math"
student2012.sub.summary2 <- summarise(group_by(student2012.sub, name), 
                                     male=length(PV1MATH[ST04Q01=="Male"])/length(PV1MATH), 
                                     female=length(PV1MATH[ST04Q01=="Female"])/length(PV1MATH))
student2012.sub.summary.m2 <- melt(student2012.sub.summary2)
colnames(student2012.sub.summary.m2)[2] <- "gender"
colnames(student2012.sub.summary.m2)[3] <- "prop"
student2012.sub.summary <- merge(student2012.sub.summary.m1, student2012.sub.summary.m2)
qplot(gender, math, data=student2012.sub.summary.m1, geom="boxplot") + theme(aspect.ratio=1)
# Overall there is a gender gap in math, perhaps 10 points, and the high scores are almost exclusively males
student2012.sub.summary$name <- factor(student2012.sub.summary$name, 
      levels=student2012.sub.summary1$name[order(student2012.sub.summary1$male)])
student2012.sub.summary$gender <- factor(student2012.sub.summary$gender, levels=c("female", "male"))
qplot(name, math, data=student2012.sub.summary, colour=gender, size=prop) + coord_flip()
# Some countries have no gender gap, a few surprises females better than males
student2012.sub.summary.gap <- summarise(group_by(student2012.sub.summary1, name), gap=male-female)
student2012.sub.summary.gapprop <- summarise(group_by(student2012.sub.summary2, name), prop=male-0.5)
student2012.sub.summary.gap <- merge(student2012.sub.summary.gap, student2012.sub.summary.gapprop)
student2012.sub.summary.gap$name <- factor(student2012.sub.summary.gap$name, 
      levels=student2012.sub.summary.gap$name[order(student2012.sub.summary.gap$gap)])
qplot(name, gap, data=student2012.sub.summary.gap, ylim=c(-30, 30), size=prop) + 
   geom_hline(yintercept=0, colour="grey80") + coord_flip()
# Gap is interesting, muslim countries with no gap is not because only few top girls measured
```
